Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"

### 2018-05-28 supplementary_material.tex
Reworked version to satisfy editorial requests

@@

\section{What to do before}

\subsection{Obtaining permission for tracking and studying participants}
A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes. To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities. These scientists then reviewed their observations and gave valuable feedback about the hack week outcomes. In order to be able to track and study participants, approval from an Internal Review Board is required. This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally do not involve human subject research. Ideally, participants should consent to be contacted at the application stage, which opens up the potential of tracking both participants and non-participants and allows for evaluation of the hack week.

\subsection{Determining your objectives}

@@

\subsection{Hack Week Location}
When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration. The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisor or department to attend. Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be a high priority when deciding on a location. Conference centers like the Lorentz Center at the University of Leiden provide funding along with the collaborative space, but have long lead times when applying for their spaces (typically 12-18 months). The same can be true for collaborative spaces within universities and active learning environments, which tend to be popular. Therefore, the search for a location should commence no later than twelve months before the workshop, ideally much earlier.
The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white-/chalkboards. A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
There is a natural tension between keeping the group together and providing physically separated break-out spaces. On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas. On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects. It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
There should be plenty of power outlets for participants to plug in their computers.
The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access to visitors from other institutions.
Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?

\section{Event Size and Duration}

@@

\subsection{Participant selection}

@@

\subsection{Funding}

@@

\subsection{Code of Conduct}

@@

\subsection{Communicating with the Outside World}

@@

\section{During a Hack Week}

@@

\subsection{Tutorials}
Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week. Unlike traditional lectures, in which more time is available to explore numerous topics in depth, tutorials aim to deliver only the most essential information while ensuring students remain engaged and interactive. Moreover, the need for an instructor to balance prepared material with spontaneous, interactive learning, while attending to community building and inclusive practices, can be difficult for even the most seasoned educator. Therefore careful planning and preparation is necessary for effective delivery of a hack week tutorial. Tutorials may also be video-taped and shared with a broader community. We have found that video recordings of tutorials are rather popular, with some of them drawing more than 1,000 views.

@@

\subsection{Break-Out Sessions}

@@

\subsection{Hack Sessions}

@@

\section{Facilitating Participation}

@@

\subsection{Impostor Syndrome}

@@

\section{What to do after}

@@

\section{Additional Figures}

Here, we present a number of additional figures related to the survey data presented in the main manuscript. The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 of the manuscript.

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_3_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_3_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_5_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q19_5_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}utbt
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_5_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q19_5_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:survey}
%\end{center}
\end{figure*}

\newpage
\bibliographystyle{aasjournal}
\bibliography{paper}

>>>>

%\newpage

\section{Materials and Methods}
\label{sec:suppmaterials}
We performed post-attendance surveys for AHW, GHW and NHW in 2016 and 2017. All surveys contained general questions about attitudes towards the workshop as well as open science and reproducibility, shared among all three surveys. The NHW and GHW surveys and the AHW 2017 survey were administered on site on the last day of the workshop; AHW participants in 2016 were e-mailed immediately after the workshop, with reminders after several weeks. All responses were collected within four weeks of the end of the workshop. The experimental procedures were approved by the Institutional Review Boards at UW, NYU and UC Berkeley. All participants gave their informed consent.

Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100\% in both years; the response rate for AHW was 71\% (35 out of 49) in 2016 and 82\% (37 out of 45) in 2017. The lower response rates for AHW can be explained by the generally lower rates of response for time-delayed survey in 2016. In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered. For the AHW surveys, we checked whether the demographics of the respondents different significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and find no significant difference for any of the groups except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees). Because we group responses in the career stage category into categories ``early-career'' (students and postdoctoral fellows), ``late-career'' (faculty, research scientists) and ``other'' (e.g. participants from outside academia), this discrepancy has no bearing on the results we report. For gender, we grouped participants into ``minority'' (women and non-binary) and ``non-minority'' (men) participants; similarly, for race and ethnicity, we also chose ``non-white'' (all but Caucasian participants) and ``white'' (Cauciasian) categories. Note that demographic data for GHW and NHW is only available for 2016, for AHW for both 2016 and 2017, and responding to any question was optional.

Participants were asked to respond to statements regarding these topics using a six-point Likert-type scale. All questions were anonymously recorded. No responses were discarded, and no pre-processing was performed on the data. We test for correlations between demographic characteristics (independent variable) and question responses (dependent variable) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript. For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}. The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.

\section{Additional Statistical Results}

Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}). We searched for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity. The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, and (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$). The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}

For no combination of question with a demographic variable, we find a $p$-value below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}. In the main text, we put these non-detections into context, quote those tests that near significance, coupled with a moderate to large effect size, as an indication where future research may be fruitful, and present the study's limitations.

\begin{table*}
%\begin{table*}[hbtp]
%\renewcommand{\arraystretch}{1.3}
\footnotesize
\caption{}
\begin{threeparttable} 
\begin{tabularx}{18cm}{p{3.6cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}}
\toprule
\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
 & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
\multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\ 
Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
\multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
\multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\ 
\bottomrule
\end{tabularx}
   \begin{tablenotes}
      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (associated letters correspond to panels in Figure 2 in the main text). Results are for hack weeks held in 2016 and 2017 for AHW and in 2016 for GHW and NHW, respectively. The number of respondents for each question is included in the table above. Additional figures showing the distributions of responses for each question are available in Section 3 of the SM below.}
      \item[\emph{a}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
      \item[\emph{c}]{Responses to the question of career stage were grouped into three responses: early-career stage; faculty and research scientists; other}
      \item[\emph{d}]{Responses to the question of gender identity were grouped with respect to representation within the field of study into two categories: non-minority; minority}
      \item[\emph{e}]{Responses to the question of racial and ethnic identity were grouped with respect to representation within the field of study into two categories: white; non-white}
\end{tablenotes}

\end{threeparttable}
\label{tab:survey}
%\end{table*}
\end{table*}

\section{Additional Figures}
\label{sec:suppfigures}
Here, we present a number of additional figures related to the survey data presented in the main manuscript. The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 above in the SM.

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_3_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr1}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_3_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr2}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr3}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q3_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr4}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr5}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr6}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_5_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr7}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q19_5_Q27_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr8}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_4_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr9}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q23_2_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr10}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q24_5_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr11}
%\end{center}
\end{figure*}

\begin{figure*}[h!]
%\begin{center}
%\begin{subfigure}[t]{\textwidth}
\centering
%\caption{}
\includegraphics[width=\textwidth]{Q19_5_Q30_stackedbars.eps}
%\end{subfigure}
\caption{}
\label{fig:corr12}
%\end{center}
\end{figure*}

\section{Practical Considerations}
\subsection{What to do before}

@@

\subsubsection{Obtaining permission for tracking and studying participants}
A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes. To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities. These scientists then reviewed their observations and gave valuable feedback about the hack week outcomes. In order to be able to track and study participants, approval from an Internal Review Board is required. This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally do not involve human subject research. Ideally, participants should consent to be contacted at the application stage, which opens up the potential of tracking both participants and non-participants and allows for evaluation of the hack week.

\subsubsection{Determining your objectives}

@@

\subsubsection{Hack Week Location}
When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration. The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisor or department to attend. Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be a high priority when deciding on a location. Conference centers like the Lorentz Center at the University of Leiden provide funding along with the collaborative space, but have long lead times when applying for their spaces (typically 12-18 months). The same can be true for collaborative spaces within universities and active learning environments, which tend to be popular. Therefore, the search for a location should commence no later than twelve months before the workshop, ideally much earlier.
The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white-/chalkboards. A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
There is a natural tension between keeping the group together and providing physically separated break-out spaces. On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas. On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects. It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
There should be plenty of power outlets for participants to plug in their computers.
The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access to visitors from other institutions.
Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?

\subsubsection{Event Size and Duration}

@@

\subsubsection{Participant selection}

@@

\subsubsection{Funding}

@@

\clearpage 
\subsubsection{Code of Conduct}

@@

\subsubsection{Communicating with the Outside World}

@@

\clearpage 
\subsection{During a Hack Week}

@@

\subsubsection{Tutorials}
Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week. Unlike traditional lectures, in which more time is available to explore numerous topics in depth, tutorials aim to deliver only the most essential information while ensuring students remain engaged and interactive. Moreover, the need for an instructor to balance prepared material with spontaneous, interactive learning, while attending to community building and inclusive practices, can be difficult for even the most seasoned educator. Therefore careful planning and preparation is necessary for effective delivery of a hack week tutorial. Tutorials may also be video-taped and shared with a broader community. We have found that video recordings of tutorials are rather popular, with some of them drawing more than 1,000 views.

@@

\subsubsection{Break-Out Sessions}

@@

\subsubsection{Hack Sessions}

@@

\subsubsection{Facilitating Participation}

@@

\subsubsection{Impostor Syndrome}

@@

\subsection{What to do after}

@@

\bibliographystyle{aasjournal}
\bibliography{paper}

### 2018-05-28 what.tex
Reworked version to satisfy editorial requests

Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. Additionally, they provide opportunities to build new collaborations and network. In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).

@@

Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning. They aim to synthesize different goals and strategies from both models: they are much more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon. Where a summer school is often organized around a framework of lectures and tutorials known in advance, hack weeks leave the majority of time unstructured. Tutorials at hack weeks often serve as an entry point into a topic for further exploration and learning, rather than an in-depth study. This is closely related to the nature of the material taught at these events. Data science is a new, emerging area combining knowledge and skills from many different fields, only a fraction of which can be communicated within a short tutorial. As organizers, we make a curated selection of a small number of tutorials each year, and encourage participants to use the unstructured time to self-organize tutorials on any topic as needed.%, often on topics vastly different to those discussed in the formal lectures.

The unstructured schedule and inherent lack of focus on specific topics within a hack week present both an opportunity and a risk. They uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number of tutorials and their content. This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities. They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few. Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach and work aimed at improving the scientific community itself.

There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week can run the danger of not doing any of them well. Because tutorials are not the major focus of a hack week, the knowledge gained by participants in these tutorials may not be very deep. Additionally, a hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require more preparation from the side of participants than other events (e.g. define project ideas, research possible connections to other participants). 
Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time (see also the following sections as well as the SM, Section 4).

>>>>

%Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. Additionally, they provide opportunities to build new collaborations and network. 
Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and building collaborations.  
In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).

@@

Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning. They aim to synthesize different goals and strategies from both models: they are more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon. Where a summer school is often organized around a framework of lectures and tutorials known in advance, hack weeks leave the majority of time to be designed by participants, under careful facilitation of the organizers. Tutorials at hack weeks often serve as an entry point into a topic for further exploration and learning. %This is closely related to the nature of the material taught at these events. Data science is a new, emerging area combining knowledge and skills from many different fields, only a fraction of which can be communicated within a short tutorial. As organizers, we make a curated selection of a small number of tutorials each year, and encourage participants to use the unstructured time to self-organize tutorials on any topic as needed.%, often on topics vastly different to those discussed in the formal lectures.

Hack weeks uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number and content of tutorials. This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities. They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few. Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach and work aimed at improving the scientific community itself.

There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week might potentially not do any of them well. Because tutorials are not necessarily the major focus of a hack week, the knowledge gained by participants in these tutorials may be shallow. A hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require significant preparation from the side of participants. 
Because of these risks, organizers face a much larger degree of uncertainty, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, and facilitation of the large range of different types of activities at any given time (see also the SM, Section 4.2.4).

### 2018-05-28 why.tex
Reworked version to satisfy editorial requests

\begin{itemize}
\item{\textit{Education and Training}:
While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component that entails discussions about reproducible research. Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices. This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}

@@

\noindent It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 1.2).

>>>>

\begin{itemize}
\item{\textit{Education and Training}:
While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning. This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}

@@

\noindent Note that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 4.1.2).
